\pkg{dataMaid}: your assistant for data cleaning in \proglang{R}

### 2017-01-30 latex/article_vol2.Rnw
Fixed

One last package that should be mentioned in this context is
\pkg{DataExplorer}. While this package does not address data cleaning
issues per se, its general strategy is quite similar to that of
\pkg{dataMaid} and to the paradigms presented below. This package
provides a few simple, but practical, tools for exploratory data
analysis, including automated documentation. Therefore, we find
\pkg{DataExplorer} to be a good candidate for a next-step package
after data cleaning is finished.

@@

We will now build four new functions and show both how they can be
used interactively and how they can be integrated with the \code{clean()}
function. These four new functions are:
\begin{description}
\item[\code{isID}] A new \code{checkFunction} intended for use in the precheck-stage. This function checks whether a variable consists exclusively of long ($> 10$ characters/digits) entries that are all of equal length, as this might be personal identification codes that we do not wish to print out in the data summary.
\item[\code{mosaicVisual}] A new \code{visualFunction} that produces mosaic plots. This function will be used in the \textit{visualize} step of \code{clean()}.
\item[\code{countZeros}] A new \code{summaryFunction} that counts the number of occurrences of the value \code{0} in a variable. This function will be used in the \textit{summarize} step of \code{clean()}.
\item[\code{identifyColons}] A new \code{checkFunction} that flags variables in which values have colons that appear before and after alphanumerical characters. This is practical for identifying autogenerated interaction effects. This function will be used in the \textit{check} step of \code{clean()}.
\end{description}
These functions are defined in turn below, and afterwards, an example of how they can be called from \code{clean()} is provided.

>>>>

One last package that should be mentioned in this context is
\pkg{DataExplorer} \citep{DataExplorer}. While this package does not
address data cleaning issues \emph{per se}, its general strategy is
quite similar to that of \pkg{dataMaid} and to the paradigms presented
below. This package provides a few simple, but practical, tools for
exploratory data analysis, including automated
documentation. Therefore, we find \pkg{DataExplorer} to be a good
candidate for a next-step package after data cleaning is finished.

@@

We will now build four new functions and show both how they can be
used interactively and how they can be integrated with the \code{clean()}
function. These four new functions are:
\begin{description}
\item[\code{isID()}] A new \code{checkFunction} intended for use in the precheck-stage. This function checks whether a variable consists exclusively of long ($> 10$ characters/digits) entries that are all of equal length, as this might be personal identification codes that we do not wish to print out in the data summary.
\item[\code{mosaicVisual()}] A new \code{visualFunction} that produces mosaic plots. This function will be used in the \textit{visualize} step of \code{clean()}.
\item[\code{countZeros()}] A new \code{summaryFunction} that counts the number of occurrences of the value \code{0} in a variable. This function will be used in the \textit{summarize} step of \code{clean()}.
\item[\code{identifyColons()}] A new \code{checkFunction} that flags variables in which values have colons that appear before and after alphanumerical characters. This is practical for identifying autogenerated interaction effects. This function will be used in the \textit{check} step of \code{clean()}.
\end{description}
These functions are defined in turn below, and afterwards, an example of how they can be called from \code{clean()} is provided.
